Thermodynamic fragments in magic state theories

### 2020-07-07 magic_paper.tex
Added content in Fragments section. Calculations seem hard, the trick doesnt work, but stabilizer fragments look kinda interesting

Of practical importance are magic state distillation (MSD) processes of the form
\begin{equation}\label{eq:distprocess}
    \rho^{\otimes k} \xrightarrow{\E \in \O_\sigma} \tau,
\end{equation}
where $k$ copies of a noisy magic state $\rho$ are converted into a target single-copy state $\tau$. \nick{expand?}

In general, the Wigner components of a $k$-copy state $\rho^{\otimes k}$ can be calculated, along with their multiplicities, from $W_\rho$ by expanding the terms in the multinomial expansion $\left( \sum_{\bm{z} \in \cal{P}_d} W_\rho(\bm{z}) \right)^n$.
This follows from the multiplicativity of the Wigner distribution.

As an example, consider the qutrit Strange state $\ketbra{\rm{S}}$ depicted in~\cref{fig:strange}. 
It is the simplest non-trivial case to analyse because the Strange state only has two distinct components $\left\{ -\frac{1}{3}, \frac{1}{6} \right\}$ with multiplicities $1$ and $8$ respectively.
Calculating the binomial expansion for the components of $\ket{S}\bra{S}^{\otimes k}$ gives $\{(-1)^j 2^{j-k} 3^{-k}\}_{0 \leq j \leq k}$ with multiplicity $8^{k-j} \binom{k}{j}$ for the $j$-th term.
This allows analytical calculation of all Lorenz curve points, and for the $k$-copy state we have that
\begin{equation}\label{eq:Skmax}
    \max_x{\lc{\ketbra{S}^{\otimes k}} {\sigma}}(x) = \frac{1}{2}\left (1 + \left(\frac{5}{3}\right)^k \right),
\end{equation}
independent of the particular $\sigma$--fragment. \ddd{[We should also give for what value of $x$ this maximum occurs.]} \nick{It will have the form of~\cref{eq:maxloc} and probably cannot be simplified more other than finding $k_\star$ exactly.}\ddd{[Finding $k_\star$ is perfect. We'll always have the complexity of the free state distribution to deal with. For the unital fragment and thermal fragment this will be computable though.]}
%1 + \left(\frac{4}{3}\right)^k \sum\limits_{j=0}^{\floor{\frac{k-1}{2}}} 4^{-(2j+1)} \binom{k}{2j+1}
%\sum_{n=1, n \mbox{ odd}}^k a^n \binom{k}{n} = \frac{1}{2} \left [  (1+a)^k - (1-a)^k \right ]

Consider the problem of purifying out magic from a noisy Strange state,
\begin{equation}
    \rho_{\rm{S}}(\epsilon) = (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \sigma,
\end{equation}
in $\O_\sigma$.
At noise levels $\epsilon \geq \frac{3}{4}$, the state is sufficiently noisy so as the Wigner distribution $\W{\rho_{\rm{S}}(\epsilon)}$ no longer contains negativities. 
At any $\epsilon \leq \frac{3}{4}$, we can investigate the distillation process resulting in a single-copy state with sufficiently low target $\epsilon_{\tau}$,
\begin{equation}\label{eq:sdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes k} \xrightarrow{\E \in \O_\sigma} \rho_{\rm{S}}(\epsilon_{\tau})
\end{equation}

The more noisy the initial state is, the more copies needed to reach the target, therefore it makes sense to define a noise level threshold 
\begin{equation}\label{eq:ethresh}
	\epsilon_{\rm{th}}(k, \epsilon_\tau, \sigma) \coloneqq \sup{\{\epsilon:\ \lc{\rho_{\rm{S}}(\epsilon)^{\otimes k}}{\sigma} \geq \lc{\rho_{\rm{S}}(\epsilon_{\tau})}{\sigma}\}},
\end{equation}
so that the MSD process~(\ref{eq:sdist}) is impossible for all $\epsilon > \epsilon_{\rm{th}}(k, \epsilon_\tau, \sigma)$

\nick{Condition in~\cref{eq:ethresh} is informally written. Can also define 
\begin{equation}\label{eq:ethresh}
	k_{\rm{th}}(\epsilon, \epsilon_\tau, \sigma) \coloneqq \min{\{k:\ k \in \mathbb{N}, \lc{\rho_{\rm{S}}(\epsilon)^{\otimes k}}{\sigma} \geq \lc{\rho_{\rm{S}}(\epsilon_{\tau})}{\sigma}\}}.
\end{equation}}
We now examine the MSD process~(\ref{eq:sdist}) in specific $\sigma$--fragments.

@@

Magic state distillation circuits in principle consist of bulk sequences of random Clifford unitaries ($\rcu$)\nick{CITE}, depicted in~\cref{fig:zoo}.
Operations in $\rcu$ can be expressed as
\begin{equation}
    \E(\rho) = \sum_i p_i U_i \rho U_i^\dagger,\ U_i \in \cd.
\end{equation}
Depending on the symmetries of such operations, a sequence of them may belong in various $\sigma$--fragments and the majorization condition~(\ref{eq:majbound}) needs to be checked in all these $\sigma$-fragments that reflect the symmetries of the operation sequence.
However, they always belong to the \emph{unital fragment} $\O_{\frac{1}{d}\id}$

@@

@@

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.5]{figs/thermal_distill.pdf}
    \caption{\textbf{Noise thresholds in thermal fragments.} Lorenz curve ratios for the Strange state purifying process in~\cref{eq:sdist} with $\epsilon_\tau = 0.05$.
    The peaks of each curve indicate the optimal temperature $\beta_{\rm{max}}$ that allows for the highest noise threshold at every given number of initial state copies $k$.
    The line $\epsilon = \frac{3}{4}$ indicates the threshold noise beyond which the Strange state no longer contains negativities.
    }
    \label{fig:thermal_distill}
\end{figure}

>>>>

Of practical importance are magic state distillation (MSD) processes of the form
\begin{equation}\label{eq:distprocess}
    \rho^{\otimes n} \xrightarrow{\E \in \O_\sigma} \tau,
\end{equation}
where $n$ copies of a noisy magic state $\rho$ are converted into a target single-copy state $\tau$. \nick{expand?}

In general, the Wigner components of a $n$-copy state $\rho^{\otimes n}$ can be calculated directly from $W_\rho$, along with their multiplicities, by expanding the terms in the multinomial expansion $\left( \sum_{\bm{z} \in \cal{P}_d} W_\rho(\bm{z}) \right)^n$. \nick{is this clear?}
This follows from the multiplicativity of the Wigner distribution.

As an example, consider the qutrit Strange state $\ketbra{\rm{S}}$ depicted in~\cref{fig:strange}. 
It is the simplest non-trivial case to analyse because its Wigner distribution consists of two distinct components.
Calculating the components of $\ket{S}\bra{S}^{\otimes n}$ allows for analytical calculation of the maximum of the $n$-copy state,
\begin{equation}\label{eq:Skmax}
    \max_x{\lc{\ketbra{S}^{\otimes n}} {\sigma}}(x) = \frac{1}{2}\left (1 + \left(\frac{5}{3}\right)^n \right).
\end{equation}
Its location $x_{k_\star}$ depends on the $\sigma$--fragment and is given by~\cref{eq:maxloc} with $k_\star = \frac{1}{2}(7^n + 9^n)$.
%1 + \left(\frac{4}{3}\right)^n \sum\limits_{j=0}^{\floor{\frac{n-1}{2}}} 4^{-(2j+1)} \binom{n}{2j+1}
%\sum_{i=j, j: odd}^n a^j \binom{n}{j} = \frac{1}{2} [ (1+a)^n - (1-a)^n ]
%$\{(-1)^j 2^{j-n} 3^{-n}\}_{0 \leq j \leq n}$ with multiplicity $8^{n-j} \binom{n}{j}$ for the $j$-th term

Consider the problem of purifying out magic from a noisy Strange state,
\begin{equation}
    \rho_{\rm{S}}(\epsilon) = (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \sigma,
\end{equation}
in $\O_\sigma$.
At noise levels $\epsilon \geq \frac{3}{4}$, the state is sufficiently noisy so as the Wigner distribution $\W{\rho_{\rm{S}}(\epsilon)}$ no longer contains negativities. 
At any $\epsilon \leq \frac{3}{4}$, we can investigate the distillation process resulting in a single-copy state with sufficiently low target $\epsilon_{\tau}$,
\begin{equation}\label{eq:sdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \xrightarrow{\E \in \O_\sigma} \rho_{\rm{S}}(\epsilon_{\tau})
\end{equation}

The more noisy the initial state is, the more copies needed to reach the target, therefore it makes sense to define a noise level threshold 
\begin{equation}\label{eq:ethresh}
	\epsilon_{\rm{th}}(n, \epsilon_\tau, \sigma) \coloneqq \sup{\{\epsilon:\ \lc{\rho_{\rm{S}}(\epsilon)^{\otimes n}}{\sigma} \geq \lc{\rho_{\rm{S}}(\epsilon_{\tau})}{\sigma}\}},
\end{equation}
so that the MSD process~(\ref{eq:sdist}) is impossible for all $\epsilon > \epsilon_{\rm{th}}(n, \epsilon_\tau, \sigma)$

\nick{Condition in~\cref{eq:ethresh} is informally written. Can also define 
\begin{equation}\label{eq:ethresh}
	n_{\rm{th}}(\epsilon, \epsilon_\tau, \sigma) \coloneqq \min{\{n:\ n \in \mathbb{N}, \lc{\rho_{\rm{S}}(\epsilon)^{\otimes n}}{\sigma} \geq \lc{\rho_{\rm{S}}(\epsilon_{\tau})}{\sigma}\}}.
\end{equation}}
We now examine the MSD process~(\ref{eq:sdist}) in specific $\sigma$--fragments.

@@

Magic state distillation circuits in principle consist of bulk sequences of random Clifford unitaries ($\rcu$)\nick{CITE}, depicted in~\cref{fig:zoo}.
Operations in $\rcu$ can be expressed as
\begin{equation}
    \E(\rho) = \sum_i p_i U_i \rho U_i^\dagger,\ U_i \in \cd.
\end{equation}
Depending on the symmetries of such operations, a sequence of them may belong in various $\sigma$--fragments and the majorization condition~(\ref{eq:majbound}) needs to be checked in all these $\sigma$-fragments that reflect the symmetries of the operation sequence.
However, they always belong to the \emph{unital fragment} $\O_{\frac{1}{d}\id}$.

@@

Lorenz curve analysis simplifies significantly in the unital fragment \nick{or so I thought initially}, because the invariant Wigner distribution is the uniform distribution depicted in~\cref{fig:maxmix}.
Therefore, $\W[\bmu]{(\frac{1}{d}\id)^{\otimes n}} = \frac{1}{d^{2n}}$ for all $\bmu \in \pd[d^n]$ and the Lorenz curve ordering only depends on $\rho$ \nick{rephrase}
The Lorenz curve $\lc{\ketbra{\rm{S}}^{\otimes n}}{\frac{1}{d}\id}$ consists of $n+1$ line segments, with $n$ endpoints, or ``elbows'', between $(0,0)$ and $(1,1)$. 
The $k_\star$-th endpoint is the maximum in~\cref{eq:Skmax} occuring at 
\begin{equation}
	x_{k_\star} = \frac{1}{2}\left(1 + \left( \frac{7}{9} \right)^n \right).
\end{equation}

All $n$ elbows, depicted in the diagram of~\cref{fig:lccalc}, are given by the following sums, $\left( x_k, \lc{\ketbra{\rm{S}^{\otimes n}}}{\frac{1}{d}\id}(k) \right) =$
\begin{equation}\label{eq:leftelb}
	\left( \frac{1}{9^n} \sum\limits_{j=0}^{k-1} 8^{2j} \binom{n}{2j}, \frac{1}{3^n} \sum\limits_{j=0}^{k-1} 4^{2j} \binom{n}{2j} \right)
\end{equation}
for all elbows left of the maximum, $k = 1, \dots, k_\star = \floor{\frac{n}{2}}+1$ and $\left( x_k, \lc{\ketbra{\rm{S}^{\otimes n}}}{\frac{1}{d}\id}(k) \right) = (x_{k_\star}, L(k_\star)) + $
\begin{equation}\label{eq:rightelb}
	\left( \left(\frac{8}{9}\right)^n \sum\limits_{j=0}^{k'-1} 8^{-(2j+1)} \binom{n}{2j+1}, \left(\frac{4}{3}\right)^n \sum\limits_{j=0}^{k'-1} 4^{-(2j+1)} \binom{n}{2j+1} \right),
\end{equation}
for all elbows right of the maximum, $k' = k_\star, \dots, n$.

\begin{figure}%
    \centering
    \subfigure[][]{%
    \label{fig:lcs_maxmix}%
    \includegraphics[height=3cm]{figs/lcs_maxmix.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_thermal}%
    \includegraphics[height=3cm]{figs/lcs_thermal.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:lcs_zero}%
    \includegraphics[height=3cm]{figs/lcs_zero.pdf}
    %\caption{Bound state}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_one}%
    \includegraphics[height=3cm]{figs/lcs_one.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Lorenz curves of Strange state copies.} Lorenz curves of $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ for $n=1,2,3,4$.
    Solid lines represent pure Strange states ($\epsilon = 0$); dashed lines represent noisy Strange states ($\epsilon = 0.1$).
    \subref{fig:lcs_maxmix} Unital fragment; \subref{fig:lcs_thermal} Thermal fragment with $H = (0,1,2)$ and $\beta = 0.5$; \subref{fig:lcs_zero} Fragment $\O_{\ketbra{0}{0}}$ (corresponds to red curve in~\cref{fig:stab_distill}); \subref{fig:lcs_one} Fragment $\O_{\ketbra{1}{1}}$ (corresponds to blue curve in~\cref{fig:stab_distill}). 
    }%
    \label{fig:wstate_examples}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/lccalcdiagram.pdf}
    \caption{\textbf{Anatomy of Strange state Lorenz curve}
    There are $n$ ``elbows'' in this Lorenz curve of an $n$-copy Strange state.
    Labels $k, k'$ indicate elbows left and right of the maximum respectively.
    Their coordinates are given in~\cref{eq:leftelb,eq:rightelb}.
    }
    \label{fig:lccalc}
\end{figure}

We plot the Lorenz curves of pure and noisy Strange state $n$-copy states for $n$ up to $4$ in~\cref{fig:lcs}.
For $n=1$, the Lorenz curve is simply additive, 
\begin{equation}
\lc{(1-\epsilon)\rho + \epsilon\sigma}{\sigma} = (1-\epsilon)\lc{\rho}{\sigma} + \epsilon\lc{\sigma}{\sigma},
\end{equation}
for any $\sigma, \rho$, but for $n\geq 2$, calculations become more complicated.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/thermal_distill.pdf}
    \caption{\textbf{Threshold dependence on temperature in thermal fragments.} Lorenz curve ratios for the Strange state purifying process in~\cref{eq:sdist} with $\epsilon_\tau = 0.05$.
    The peaks of each curve indicate the optimal temperature $\beta_{\rm{max}}$ that allows for the highest noise threshold at every given number of initial state copies $k$.
    The line $\epsilon = \frac{3}{4}$ indicates the threshold noise beyond which the Strange state no longer contains negativities.
    }
    \label{fig:thermal_distill}
\end{figure}

\subsection{Magic in pure stabilizer fragments}\label{sec:magstab}

There exist $d(d+1)$ pure stabilizer states for a $d$-dimensional system, all of which correspond to a uniform distribion over an affine line in the phase space \nick{CITE}.
As an example, the distribution of the qutrit zero state $\ketbra{0}$ is illustrated in~\cref{fig:zero}.
Any pure stabilizer state $\sigma$ is a rank--$1$ state and contains $d^2 - d$ zeros in its Wigner distribution, so we approximate it by the full--rank state $\sigma' = (1-\epsilon)\sigma + \epsilon \frac{1}{d}\id$ for some infinitesimal $\epsilon > 0$.

We explore the distillation process of~\cref{eq:sdist} in these modified fragments $\O_\sigma'$ in~\cref{fig:stab_distill}.
We repeat the plot for distillation of Norrell states in~\cref{fig:stab_distill_norrell}.

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/stab_distill.pdf}
    \caption{\textbf{Strange state distillation thresholds in stabilizer fragments.} Lorenz curve thresholds for the Strange state purifying process in~\cref{eq:sdist} with $\epsilon_\tau = 0.05$.
    All $12$ stabilizer fragments produce two distinct trends: ($k=0$ - red curve) the stabilizer states that have non-zero value at the origin of the phase space lie low and ($k=1,2$ - blue curve) the stabilizer states that have a zero value at the origin of the phase space lie high.
    This is specific to the Strange state.
    For high enough copies, these Lorenz curves surpass the unital noise threshold $\epsilon_{\rm{th}} = \frac{3}{4}$.
    }
    \label{fig:stab_distill}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/stab_distill_norrell.pdf}
    \caption{\textbf{Norrell state distillation thresholds in stabilizer fragments.} Lorenz curve thresholds for the Strange state purifying process in~\cref{eq:sdist} with $\epsilon_\tau = 0.05$.
    For high enough copies, certain Lorenz curves surpass the unital noise threshold $\epsilon_{\rm{th}} = \frac{3}{4}$.
    }
    \label{fig:stab_distill_norrell}
\end{figure}
